---
title: "stats_code"
author: "Aimee"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(readxl)
library(knitr)  
setwd("C:/Users/aimee/OneDrive - University of Cape Town/OneDrive - University of Cape Town/msc_thesis/parechinus/data")
```

#Survival rate
```{r}
survival <- read_excel("C:/Users/aimee/OneDrive - University of Cape Town/OneDrive - University of Cape Town/msc_thesis/parechinus/data/survival.xlsx")

survival$Treatment <- paste0(substr(survival$Diet, 1, 1), substr(survival$Temperature, 1, 1))

survival <- survival %>% group_by(Treatment) %>% mutate(CumulativeLost= cumsum(Lost))

survival <-survival %>% group_by(Treatment) %>% mutate(survival_rate = 100*( (76-CumulativeLost)/76))


```

Check normality and homoscedasticity

For Shapiro: If p < 0.05, we can be 95% certain that the 
data are not normally distributed. 

For Levene: If p < 0.05, the treatment group data does not have equal variances
```{r}

final_survival <- survival %>% filter(Week == 24)
week8_survival <- survival%>% filter(Week == 8)

shapiro.test(week8_survival$survival_rate) #not normal
#W = 0.82452, p-value = 0.0001221


shapiro.test(final_survival$survival_rate) #not normal
#W = 0.8611, p-value = 0.003538

shapiro.test(survival$survival_rate) #not normal
#W = 0.82159, p-value < 2.2e-16

require("car")

leveneTest(week8_survival$survival_rate~week8_survival$Treatment) #equal
#      Df F value  Pr(>F)  
#group  7  2.0566 0.08893 .
#      24             

leveneTest(final_survival$survival_rate~final_survival$Treatment) #equal
#      Df F value Pr(>F)
#group  5  1.6273 0.2035
#      18         

```


Kruskal-Wallis for week 8 survival
```{r}
#install.packages("dunn.test", dependencies = TRUE)
require("dunn.test")
#install.packages("coin", dependencies = TRUE)
require(coin)

kruskal.test(survival_rate ~ Treatment, data = week8_survival) 
#significant: Kruskal-Wallis chi-squared = 18.163, df = 7, p-value = 0.01126

dunn.test(week8_survival$survival_rate, g = week8_survival$Treatment)


```


Kruskal-Wallis for final survival
```{r}


kruskal.test(survival_rate ~ Treatment, data = final_survival) #not significant: Kruskal-Wallis chi-squared = 9.9497, df = 5, p-value = 0.07667



```



#SGR

Load data
```{r}
size <- read_excel("C:/Users/aimee/OneDrive - University of Cape Town/OneDrive - University of Cape Town/msc_thesis/parechinus/data/parechinus_size.xlsx")


weight <- read_excel("C:/Users/aimee/OneDrive - University of Cape Town/OneDrive - University of Cape Town/msc_thesis/parechinus/data/parechinus_weight.xlsx")

somatic_growth <- read_excel("C:/Users/aimee/OneDrive - University of Cape Town/OneDrive - University of Cape Town/msc_thesis/parechinus/data/somatic_growth.xlsx")
```

Check normality and homoscedasticity:

Actual data values
```{r}
shapiro.test(size$diam) #not normal
#W = 0.99446, p-value = 4.329e-09

shapiro.test(weight$`weight (g)`) #not normal
#W = 0.99001, p-value = 1.749e-13

require("car")

leveneTest(size$diam~size$treatment) #not equal p=7.913e-10 ***

leveneTest(weight$`weight (g)`~weight$treatment) #not equal p= 2.937e-07 ***

#not normal and variances not equal - so Kruskal Wallis
```
SGR values at weekly intervals
Normality & Homoscedasticity tests
```{r}

weight_SGR_prior_tests <- function(data){
  shapiro <- shapiro.test(data$`weight_specific growth rate`)
  levene <- leveneTest(data$`weight_specific growth rate`~data$treatment)
  print(shapiro)
  print(levene)
}

size_SGR_prior_tests <- function(data){
  shapiro <- shapiro.test(data$`size_specific growth rate`)
  levene <- leveneTest(data$`size_specific growth rate`~data$treatment)
  print(shapiro)
  print(levene)
}

SGR4 <- somatic_growth %>% filter(week == 4)
SGR8 <- somatic_growth %>% filter(week == 8)
SGR13 <- somatic_growth %>% filter(week == 13)
SGR18 <- somatic_growth %>% filter(week == 18)
SGR23 <- somatic_growth %>% filter(week == 23)
  
weight_SGR_prior_tests(SGR4) #normal & equal - ANOVA
weight_SGR_prior_tests(SGR8) #normal & NOT equal - welch ANOVA
weight_SGR_prior_tests(SGR13) #normal & equal - ANOVA
weight_SGR_prior_tests(SGR18)#normal & equal - ANOVA
weight_SGR_prior_tests(SGR23)#normal & equal - ANOVA

weight_SGR_prior_tests(somatic_growth) #normal & NOT equal
 
size_SGR_prior_tests(SGR4)#NOT normal & equal - Kruskal Wallis
size_SGR_prior_tests(SGR8) #normal & equal - ANOVA
size_SGR_prior_tests(SGR13) #normal & equal - ANOVA
size_SGR_prior_tests(SGR18) #normal & equal - ANOVA
size_SGR_prior_tests(SGR23) #normal & equal - ANOVA

```

ANOVA tests
```{r}
anova_weight_SGR <- function(dat){
aov(`weight_specific growth rate`~diet*temperature,data=dat)
}
w_anova4<- anova_weight_SGR(SGR4)
w_anova13<- anova_weight_SGR(SGR13)
w_anova18<- anova_weight_SGR(SGR18)
w_anova23<- anova_weight_SGR(SGR23)
summary(w_anova4)

summary(w_anova13)
summary(w_anova18)
summary(w_anova23)

anova_size_SGR <- function(dat){
aov(`size_specific growth rate`~diet*temperature,data=dat)
}

kruskal.test(`size_specific growth rate` ~ treatment, data = SGR4)
s_anova8<- anova_size_SGR(SGR8)
s_anova13<- anova_size_SGR(SGR13)
s_anova18<- anova_size_SGR(SGR18)
s_anova23<- anova_size_SGR(SGR23)
summary(s_anova8)
summary(s_anova13)
summary(s_anova18)
summary(s_anova23)



```



Post hoc
```{r}
Tukey4 = TukeyHSD(w_anova4, ordered=TRUE) 
Tukey4
```



Averages
```{r}
SGR4 %>% group_by(treatment) %>% summarise(SGR_mean = mean(`weight_specific growth rate`), SGR_se = sd(`weight_specific growth rate`)/sqrt(length(`weight_specific growth rate`)))




```


#Gonad

Load data
```{r}
gonad <- read_excel("C:/Users/aimee/OneDrive - University of Cape Town/OneDrive - University of Cape Town/msc_thesis/parechinus/data/gonad.xlsx")

gonad9 <- gonad %>% filter(week == 9)
gonad19 <- gonad %>% filter(week == 19)
gonad26 <- gonad %>% filter(week == 26)


```

Normality & Homoscedasticity
```{r}
require(car)

GSI_prior_tests <- function(data){
  shapiro <- shapiro.test(data$GSI)
  levene <- leveneTest(data$GSI~data$treatment)
  print(shapiro)
  print(levene)
}

GSI_prior_tests(gonad9) #not normal, equal variance - Kruskal
GSI_prior_tests(gonad19) #normal, equal variance - ANOVA
GSI_prior_tests(gonad26) #normal, equal variance - ANOVA


```
Kruskals
```{r}


library(dunn.test)
kruskal.test(GSI ~ treatment, data = gonad9) #significant
#Kruskal-Wallis chi-squared = 32.586, df = 7, p-value = 3.162e-05

require("dunn.test")
#install.packages("coin", dependencies = TRUE)
require(coin)

##DIET##
kruskal.test(GSI ~ diet, data = gonad9) #significant
#Kruskal-Wallis chi-squared = 31.198, df = 3, p-value =7.724e-07

dunn.test(gonad9$GSI, g = gonad9$diet)
dunn.test(gonad9$GSI, g = gonad9$treatment)
##TEMP##
kruskal.test(GSI ~ temperature, data = gonad9) #not significant
#Kruskal-Wallis chi-squared = 0.132, df = 1, p-value = 0.7164




dunn.test(gonad9$GSI, g = gonad9$treatment)

dunn.test(week8_survival$survival_rate, g = week8_survival$Treatment)
```

ANOVA
```{r}
anova_GSI <- function(dat){
aov(GSI~diet,data=dat)
}

anova19 <- anova_GSI(gonad19)
summary(anova19)

Tukey19 = TukeyHSD(anova19, ordered=TRUE) 
Tukey19

anova26 <- anova_GSI(gonad26)
summary(anova26)

Tukey26 = TukeyHSD(anova26, ordered=TRUE) 
Tukey26
```

